\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Type-Preserving Mutations

### 2021-04-16 papers/oopsla-21/main.tex
NIT

%Instead of continuously generating random invalid test cases from scratch,
\fuzzchick keeps a queue of interesting previously executed test cases that can
be mutated using high-level, type-preserving transformations to produce new
ones.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a much higher probability of producing a new interesting
test case than generating a new one from scratch using a na\"ive generator.
%
In this light, \fuzzchick is likely to preserve the semantic structure of the
mutated data, as mutations are applied directly at the data type level, e.g.,
the program's AST in case of generating code. Internally, \fuzzchick relies on
execution traces to distinguish which executed test cases were interesting and
are therefore worth mutating --- a popular technique known as
\emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
\cite{afl}.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces --- any other test case is simply thrown away.
%
The techniques behind \fuzzchick are not meant to replace sophisticated,
manually written random generators, but rather providing an adequate testing
solution for early stages of development.
%, and while a better test suite using manually-written generators is
%still under construction.

@@

%
%% Problems of FuzzChick: mutations are superficial, long queues, power
%% scheduler
%
%% In this work, we take a close look at \fuzzchick for opportunities to improve.
%
%% Notably, we establish several aspects of \fuzzchick that can be
%% revised.%
%
The approach taken by \fuzzchick leaves considerable room for improvements in
many aspects\footnote{Some of these were already identified by the authors of
  \fuzzchick in their original work.}:
%But \fuzzchick suffers several disadvantages.%
%
In particular:
%
\begin{inparaenum}
\item type-preserving mutations, if not done carefully, can become superficial
  and rely heavily on randomness to produce diverse mutants,
\item the queuing mechanism can cause large delays if interesting values are
  enqueued frequently and there is no way to prioritize them, and
\item using an heuristic to assign a certain ``mutation energy'' to each
  interesting test case (often referred to as a \emph{power schedule}) requires
  fine tuning and can be hard to generalize to work well under different testing
  scenarios.
\end{inparaenum}
%
Moreover, we replicated the Information-Flow Control (IFC) stack machine case
study bundled with \fuzzchick and observed a surprising limitation:
%
after repeating each experiment 30 times, \fuzzchick could only find 5 out of
the 20 injected bugs with a 100\% efficacy.
%
%% the hardest one being found only around 13\% of the time after an hour of
%% testing.
%
Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
better than using normal random testing coupled with na\"ive random generators
by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
%
However, we believe that a more meticulous evaluation ought to take failure to
find a counterexample as an important metric when comparing PBT tools, not just
mean time to failure --- when one is found.

%% Our solution
In this work, we introduce \mutagen, a PBT tool implemented %in Haskell
that also builds upon code instrumentation and type-preserving mutations (see
Section \ref{sec:mutagen}).
%
% the basic ideas behind \fuzzchick, i.e.,
%
%% \mutagen is introduced in detail in Section \ref{sec:mutagen}.
%
%% Exhaustive mutations
Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
%
That is, given an interesting, previously executed test case, our tool
precomputes and schedules every type-preserving mutation that can be applied to
it.
%
This approach has two inherent advantages. %when compared to \fuzzchick.
%
On one hand, mutations do not rely on randomness to be generated.
%
Instead, every subexpression of the input test case is mutated on the same
basis, which guarantees that no interesting mutation is omitted due to
randomness.
%
%% On the other hand, scheduling mutations exhaustively eliminates the need for a
%% power schedule to assign a given mutation energy to each input test case.
%
On the other hand, scheduling mutations exhaustively eliminates the need for a
power schedule to assign a given mutation energy to each input test case.

>>>>

%Instead of continuously generating random invalid test cases from scratch,
\fuzzchick keeps a queue of interesting previously executed test cases that can
be mutated using high-level, type-preserving transformations to produce new
ones.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a much higher probability of producing a new interesting
test case than generating a new one from scratch using a na\"ive generator.
%
In this light, \fuzzchick is likely to preserve the semantic structure of the
mutated data, as mutations are applied directly at the data type level, e.g.,
the program's AST in case of generating code.
%
Internally, \fuzzchick relies on execution traces to distinguish which executed
test cases were interesting and are therefore worth mutating --- a popular
technique known as \emph{coverage-guided fuzzing} and popularized by tools like
\emph{AFL} \cite{afl}.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces --- any other test case is simply thrown away.
%
The techniques behind \fuzzchick are not meant to replace sophisticated,
manually written random generators, but rather to provide an adequate testing
solution for early stages of development.
%, and while a better test suite using manually-written generators is
%still under construction.

@@

%
%% Problems of FuzzChick: mutations are superficial, long queues, power
%% scheduler
%
%% In this work, we take a close look at \fuzzchick for opportunities to improve.
%
%% Notably, we establish several aspects of \fuzzchick that can be
%% revised.%
%
The approach taken by \fuzzchick leaves considerable room for
improvements in several aspects.%
\footnote{Some of these were already identified by the authors of \fuzzchick in
  their original work.}
%But \fuzzchick suffers several disadvantages.%
%
In particular:
%
\begin{inparaenum}
\item type-preserving mutations, if not done carefully, can become superficial
  and rely heavily on randomness to produce diverse mutants,
\item the queuing mechanism can cause large delays if interesting values are
  enqueued frequently and there is no way to prioritize them, and
\item using an heuristic to assign a certain ``mutation energy'' to each
  interesting test case (often referred to as a \emph{power schedule}) requires
  fine tuning and can be hard to generalize to work well under different testing
  scenarios.
\end{inparaenum}
%
Moreover, we replicated the Information-Flow Control (IFC) stack machine case
study bundled with \fuzzchick and observed a surprising limitation:
%
after repeating each experiment 30 times, \fuzzchick could only find 5 out of
the 20 injected bugs with a 100\% efficacy.
%
%% the hardest one being found only around 13\% of the time after an hour of
%% testing.
%
Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
better than using normal random testing coupled with na\"ive random generators
by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
%
However, we believe that a more meticulous evaluation ought to take failure to
find a counterexample as an important metric when comparing PBT tools, not just
mean time to failure --- when one is found.

%% Our solution
In this work, we introduce \mutagen, a PBT tool %implemented in Haskell
that also builds upon code instrumentation and type-preserving mutations (see
Section \ref{sec:mutagen}).
%
% the basic ideas behind \fuzzchick, i.e.,
%
%% \mutagen is introduced in detail in Section \ref{sec:mutagen}.
%
%% Exhaustive mutations
Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
%
That is, given an interesting, previously executed test case, our tool
precomputes and schedules every type-preserving mutation that can be applied to
it.
%
This approach has two inherent advantages. %when compared to \fuzzchick.
%
On one hand, mutations do not rely on randomness to be generated.
%
Instead, every subexpression of the input test case is mutated on the same
basis, which guarantees that no interesting mutation is omitted due to
randomness.
%
%% On the other hand, scheduling mutations exhaustively eliminates the need for a
%% power schedule to assign a given mutation energy to each input test case.
%
On the other hand, scheduling mutations exhaustively eliminates the need for a
power schedule to assign a given mutation energy to each input test case.
